#
#Copyright (c) 1996 - 2002 FreeBSD Project
#Copyright (c) 1991, 1993
#The Regents of the University of California.  All rights reserved.
#
#Redistribution and use in source and binary forms, with or without
#modification, are permitted provided that the following conditions
#are met:
#1. Redistributions of source code must retain the above copyright
#notice, this list of conditions and the following disclaimer.
#2. Redistributions in binary form must reproduce the above copyright
#notice, this list of conditions and the following disclaimer in the
#documentation and/or other materials provided with the distribution.
#4. Neither the name of the University nor the names of its contributors
#may be used to endorse or promote products derived from this software
#without specific prior written permission.
#
#THIS SOFTWARE IS PROVIDED BY THE REGENTS AND CONTRIBUTORS ``AS IS'' AND
#ANY EXPRESS OR IMPLIED WARRANTIES, INCLUDING, BUT NOT LIMITED TO, THE
#IMPLIED WARRANTIES OF MERCHANTABILITY AND FITNESS FOR A PARTICULAR PURPOSE
#ARE DISCLAIMED.  IN NO EVENT SHALL THE REGENTS OR CONTRIBUTORS BE LIABLE
#FOR ANY DIRECT, INDIRECT, INCIDENTAL, SPECIAL, EXEMPLARY, OR CONSEQUENTIAL
#DAMAGES (INCLUDING, BUT NOT LIMITED TO, PROCUREMENT OF SUBSTITUTE GOODS
#OR SERVICES; LOSS OF USE, DATA, OR PROFITS; OR BUSINESS INTERRUPTION)
#HOWEVER CAUSED AND ON ANY THEORY OF LIABILITY, WHETHER IN CONTRACT, STRICT
#LIABILITY, OR TORT (INCLUDING NEGLIGENCE OR OTHERWISE) ARISING IN ANY WAY
#OUT OF THE USE OF THIS SOFTWARE, EVEN IF ADVISED OF THE POSSIBILITY OF
#SUCH DAMAGE.
#
@node Locale
@chapter Locale (@file{locale.h})

A @dfn{locale} is the name for a collection of parameters (affecting
collating sequences and formatting conventions) that may be different
depending on location or culture.  The @code{"C"} locale is the only
one defined in the ANSI C standard.

This is a minimal implementation, supporting only the required @code{"C"}
value for locale; strings representing other locales are not
honored.  (@code{""} is also accepted; it represents the default locale
for an implementation, here equivalent to @code{"C"}).


@file{locale.h} defines the structure @code{lconv} to collect the
information on a locale, with the following fields:

@table @code
@item char *decimal_point
The decimal point character used to format ``ordinary'' numbers (all
numbers except those referring to amounts of money).  @code{"."} in the
C locale. 

@item char *thousands_sep
The character (if any) used to separate groups of digits, when
formatting ordinary numbers.
@code{""} in the C locale.

@item char *grouping
Specifications for how many digits to group (if any grouping is done at
all) when formatting ordinary numbers.  The @emph{numeric value} of each
character in the string represents the number of digits for the next
group, and a value of @code{0} (that is, the string's trailing
@code{NULL}) means to continue grouping digits using the last value
specified.  Use @code{CHAR_MAX} to indicate that no further grouping is
desired.  @code{""} in the C locale. 

@item char *int_curr_symbol
The international currency symbol (first three characters), if any, and
the character used to separate it from numbers.
@code{""} in the C locale.

@item char *currency_symbol
The local currency symbol, if any.
@code{""} in the C locale.

@item char *mon_decimal_point
The symbol used to delimit fractions in amounts of money.
@code{""} in the C locale.

@item char *mon_thousands_sep
Similar to @code{thousands_sep}, but used for amounts of money.
@code{""} in the C locale.

@item char *mon_grouping
Similar to @code{grouping}, but used for amounts of money.
@code{""} in the C locale.

@item char *positive_sign
A string to flag positive amounts of money when formatting.
@code{""} in the C locale.

@item char *negative_sign
A string to flag negative amounts of money when formatting.
@code{""} in the C locale.

@item char int_frac_digits
The number of digits to display when formatting amounts of money to
international conventions.
@code{CHAR_MAX} (the largest number representable as a @code{char}) in
the C locale. 

@item char frac_digits
The number of digits to display when formatting amounts of money to
local conventions.
@code{CHAR_MAX} in the C locale. 

@item char p_cs_precedes
@code{1} indicates the local currency symbol is used before a
@emph{positive or zero} formatted amount of money; @code{0} indicates
the currency symbol is placed after the formatted number.
@code{CHAR_MAX} in the C locale. 

@item char p_sep_by_space
@code{1} indicates the local currency symbol must be separated from
@emph{positive or zero} numbers by a space; @code{0} indicates that it
is immediately adjacent to numbers.
@code{CHAR_MAX} in the C locale. 

@item char n_cs_precedes
@code{1} indicates the local currency symbol is used before a
@emph{negative} formatted amount of money; @code{0} indicates
the currency symbol is placed after the formatted number.
@code{CHAR_MAX} in the C locale. 

@item char n_sep_by_space
@code{1} indicates the local currency symbol must be separated from
@emph{negative} numbers by a space; @code{0} indicates that it
is immediately adjacent to numbers.
@code{CHAR_MAX} in the C locale. 

@item char p_sign_posn
Controls the position of the @emph{positive} sign for
numbers representing money.  @code{0} means parentheses surround the
number; @code{1} means the sign is placed before both the number and the
currency symbol; @code{2} means the sign is placed after both the number
and the currency symbol; @code{3} means the sign is placed just before
the currency symbol; and @code{4} means the sign is placed just after
the currency symbol.
@code{CHAR_MAX} in the C locale. 

@item char n_sign_posn
Controls the position of the @emph{negative} sign for numbers
representing money, using the same rules as @code{p_sign_posn}.
@code{CHAR_MAX} in the C locale. 
@end table

@menu
* Function setlocale::  Select or query locale
@end menu

@page
@include locale/locale.def
